### 2019-12-09 main.tex
Abouthalf wayt through the draft with my comments

\documentclass[aps,twocolumn,showpacs,groupedaddress, nofootinbib]{revtex4}  % for review and submission
%\documentclass[aps,preprint,showpacs,superscriptaddress,groupedaddress]{revtex4}  % for double-spaced preprint
%\usepackage{graphicx}  % needed for figures
\usepackage{graphicx} % omit 'demo' for real document
\usepackage{dcolumn}   % needed for some tables
\usepackage{bm}        % for math
\usepackage{amssymb}   % for math
\usepackage{acronym}   % for acronym
\usepackage{float}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{latexsym}
\usepackage{url}
\usepackage{multirow}
\usepackage[flushleft]{threeparttable}
\usepackage{natbib}
%\usepackage[natbib, maxcitenames=3]{biblatex}
%\usepackage[maxnames=3]{biblatex}
\usepackage{subfigure}
\usepackage{amsmath}
%\usepackage[font=small,labelfont=bf,justification=justified,format=plain]{caption}
%\usepackage{booktabs, siunitx}

%\usepackage[numbers,sort]{natbib}
\bibliographystyle{unsrt3}
%\usepackage{subcaption}
%\captionsetup[figure]{slc=off} % "slc" is an abbreviation for "singlelinecheck"

% avoids incorrect hyphenation, added Nov/08 by SSR
%\hyphenation{ALPGEN}
%\hyphenation{EVTGEN}
%\hyphenation{PYTHIA}
%\renewcommand{\bibfont}{\normalfont\small}
% test comment

%\renewcommand*{\footnote}{\arabic{footnote}}

%\renewcommand{\thempfootnote}{\arabic{mpfootnote}}

\newcommand{\dcc}{LIGO-P1700428}
\newcommand{\cm}[1]{\textbf{\textcolor{red}{CM: #1}}}
\newcommand{\MC}[1]{\textcolor{red}{MC: #1}}

\def\ra{\ensuremath\rightarrow}

%% ----- input git-version tag
%\input{tag.tex}

\begin{document}

% The following information is for internal review, please remove them for submission
\widetext
%\leftline{Version xx as of \today}

% the following line is for submission, including submission to the arXiv!!
%\hspace{5.2in} \mbox{Fermilab-Pub-04/xxx-E}

\title{Detection and Classification of Supernova Gravitational Waves Signals: A Deep Learning Approach}
%\input author_list.tex       % D0 authors (remove the first 3 lines
                             % of this file prior to submission, they
                             % contain a time stamp for the authorlist)
                             % (includes institutions and visitors)
%\author{Man Leong Chan$^1$, Chris Messenger$^1$, Ik Siong Heng$^1$, Martin Hendry$^1$ $\&$ Yan Bei Chen$^2$}
%\affiliation{[1.0\linewidth]$^1$SUPA, School of Physics and Astronomy, University of Glasgow, Glasgow G12 8QQ, UK\\
%$^2$ Caltech, Pasadena, California 91125, USA}
 %$^2$ TianQin Research Center for Gravitational Physics, Sun Yat-sen University, China\\
 %$^3$ Tsinghua University, Beijing, China\\
 %$^4$ University of Western Australia, Australia\\
 %$^5$ Caltech, California, United States}

\author{Authors$^{12}$\cm{fix}}
\affiliation{
$^1$Department of Applied Physics, Fukuoka University, Nanakuma 8-19-1, Fukuoka 814-0180, Japan\\
$^2$SUPA, School of Physics and Astronomy, University of Glasgow, Glasgow G12 8QQ, UK}
%$^2$Caltech CaRT, Pasadena, California 91125, USA}
%\date{\today}
%\date{\commitDATE\\\mbox{\small \commitID}\\\mbox{\dcc}}

\begin{abstract}
We demonstrate the application of a convolutional neural network to the
gravitational wave signals from core collapse supernovae. Using simulated time
series of gravitational wave detectors, we show that based on the explosion
mechanisms, a convolutional neural network can be used to detect and classify
the gravitational wave signals buried in noise. For the waveforms used in the
training of the convolutional neural network, our results suggest that a
network of advanced LIGO, advanced VIRGO and KAGRA, or a network of LIGO A+,
advanced VIRGO and KAGRA is likely to detect a magnetorotational core collapse
supernovae within the Large and Small Magellanic Clouds, or a Galactic event if
the explosion mechanism is the neutrino-driven mechanism.  By testing the
convolutional neural network with waveforms not used for training, we show that
the true alarm probabilities are $52\%$ and $83\%$ at $60$kpc for waveforms
$\text{R3E1AC}$ and $\text{R4E1FC\_L}$.  For waveforms $\text{s}20$ and
$\text{SFHx}$ at $10$ kpc, the true alarm probabilities are $70\%$ and $93\%$
respectively. All at false alarm probability equal to $0.1$\cm{why is this nopt
also written as a percentage? Should it be 10\%?}.  \end{abstract}

\pacs{}
\maketitle
\acrodef{GW}[GW]{gravitational wave}
\acrodef{BNS}[BNS]{binary neutron star}
\acrodef{BBH}[BBH]{binary black hole}
\acrodef{NSBH}[NSBH]{neutron star black hole}
\acrodef{EM}[EM]{electromagnetic}
\acrodef{CBC}[CBC]{compact binary coalescence}
\acrodef{CNN}[CNN]{Convolutional neural network}
\acrodef{CCSN}[CCSN]{core collapse supernova}
\acrodefplural{CCSN}[CCSNe]{core collapse supernovae}
\acrodef{ROC}[ROC]{receiver operator characteristic}
\acrodef{TAP}[TAP]{true alarm probability}
\acrodef{FAP}[FAP]{false alarm probability}
\acrodef{aLIGO}[aLIGO]{advanced LIGO}
\acrodef{AdVirgo}[AdVirgo]{advanced VIRGO}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%
% general background on GWs
%
Since 2015 when LIGO made the first direct observation of \acp{GW} from the
merger of a binary black hole~\cite{abbott2016observation}, there have been
numerous observations of \acp{GW} from similar systems in the subsequent
observation runs of LIGO and VIRGO~\cite{abbott2016gw151226,
abbott2017gw170608, abbott2017gw170814}. These discoveries represent a crucial
milestone in \ac{GW} astronomy and have opened up a new window on the universe.
More recently, LIGO and VIRGO have observed the \acp{GW} from a binary neutron
star merger~\cite{abbott2017gw170817, abbott2017gravitational, abbott2017multi}
where the \acp{GW} and an associated gamma-ray burst were observed
simultaneously. Other counterparts across the electromagnetic spectrum were
also observed by later follow-up observations~\cite{abbott2017multi}. In the
near future, many more observations of \acp{GW} from similar \ac{CBC} systems
can be expected as KAGRA starts joint observations with LIGO and
VIRGO~\cite{aso2013interferometer, somiya2012detector, abbott2018prospects}.

%
% intro to core-collapse supernovae
%
In addition to \acp{CBC}, massive stars with $10-100 \text{M}_\odot $ at
zeros-age main sequence ending their lives by becoming 
%with core of $10-100 \text{M}_\odot $ prior to collapse 
\acp{CCSN} are also considered to be potential sources to the second generation
detectors such as \ac{aLIGO}~\cite{aasi2015advanced},
\ac{AdVirgo}~\cite{acernese2014advanced} and KAGRA
interferometers~\cite{aso2013interferometer, gossan2016observing,
abbott2016first}. It is currently not entirely clear to astronomers how such
massive stars become supernovae. The basic theory of the explosion, confirmed
by the neutrino events observed from SN1987A~\cite{sato1987analysis}, begins
with a massive star at the final stage of its life forming a core that is
composed of iron nuclei after it has burned all its stellar fuel via fusion
reactions. The iron core is supported by the pressure of relativistic
degenerate electrons and if the mass of the core exceeds the effective
Chandrasekhar mass~\cite{baron1990effect, bethe1990supernova}, core collapse
will ensue and continue until the core reaches nuclear density. The nuclear
equation of state will then stiffens by the strong nuclear force above the
nuclear density and stops the core collapse. The inner core will bounce back
and a shock wave will be sent through the infalling matter.  By losing energy
to the dissociation of the iron nuclei and to neutrino cooling, the shock wave
will stall. For the star to become a supernova, the shock wave will need to be
revived~\cite{o2011black}. However the mechanism via which this occurs and the
supernova explosion is caused has been the subject of intense study and remains
an unsolved problem.

%
% discussion of the CCSN mechanisms
%
There exist two most popular theories~\cm{theories for what?}, the
neutrino-driven mechanism~\cite{bethe1985revival, bethe1990supernova} and the
magnetorotational mechanism~\cite{janka2012explosion, kotake2012core,
mezzacappa2014two}. For supernova progenitors with core rotation too slow to
affect the dynamics~\cite{takiwaki2016three, summa2018rotation}, the
neutrino-driven mechanism is believed to be the active mechanism. The majority
of the observed \acp{CCSN} can be explained by the neutrino
mechanism~\cite{bruenn2016development}~\cm{why?}. The neutrino
mechanism~\cite{bethe1985revival, janka2007theory} suggests that approximately
$5-10\%$ of the outgoing neutrino luminosity is stored below the shock, which
causes turbulence to occur and thermal pressure to increase. The stalled shock
can be revived by their combined effects~\cite{couch2015role}. Producing a
\ac{CCSN} via the neutrino mechanism may also require convection and the
standing accretion shock instability~\cite{blondin2003stability}. On the other
hand, the magnetorotational mechanism requires rapid core spin and a strong
magnetic field~\cite{leblanc1970numerical, burrows2007simulations,
takiwaki2009special,
moiseenko2006magnetorotational,mosta2014magnetorotational}. Together, they may
produce an outflow that may cause the most energetic \acp{CCSN}
observed~\cm{are you referring to a real observed CCSN, if so you should
provide a reference.}. The magnetorotational mechanism may \cm{also?} be able
to explain the extreme hypernovae and the observed long gamma-ray
bursts\cite{woosley2006progenitor, yoon2005evolution, de2013rotation}.

%
% GWS from CCSN and the existing literature
%
Correctly classifying the \ac{GW} signals from a \ac{CCSN} is important in
understanding the explosion mechanism.
%Electromagnetic observations provide only indirect informations about the
%mechanism of a \ac{CCSN} because photons interact strongly with matter and
%only information from the optically thin regions far from the central core
%could reach us.  However, 
As \acp{GW} are emitted\cm{generated?} in the central core of a \ac{CCSN}, they
are likely to carry direct information of the \ac{CCSN} and therefore provide a
probe of the explosion mechanism that produces them. In \ac{GW} astronomy, for
\ac{CBC} events the established method for searching for signals is
matched-filtering~\cm{give ref for pycbc and gstlal}. However, since the
emission process of the \acp{GW} from \acp{CCSN} is affected by turbulence in
the post-bounce and is \cm{expected to be?} stochastic in nature,  the signal
evolution cannot be predicted robustly~\cite{ott2009gravitational,
kotake2013multiple, kotake2009stochastic}.  This in turn prevents
matched-filtering from being applied to \acp{CCSN}.
%Although, Neutrino radiation-hydrodynamics simulations have been an ongoing
%effort to decode and understand explosion mechanism of \acp{CCSN} and the
%\acp{GW} they produce\cite{ott2009gravitational}.
Methods and algorithms have been developed for the detection and classification
of signals from \acp{CCSN}.  For example, a method known as principle component
analysis has been developed~\cite{heng2009rotating, rover2009bayesian,
powell2015classification, powell2017classification,
suvorova2019reconstructing}. This method creates a set of component basis
vectors from a \cm{larger?} set of \ac{CCSN} waveforms belonging to a
particular mechanism where the basis vectors represent the common features of
those waveforms.  There have been other approaches developed in the literature
such as Bayesian inference~\cite{rover2009bayesian}, Bayesian model
selection~\cite{logue2012inferring} , multivariate regression
modelling~\cite{engels2014multivariate}, maximum
entropy~\cite{summerscales2008maximum}, maximum
likelihood~\cite{gursel1989near} and Tikhonov regularization
scheme~\cite{rakhmanov2006rank, hayama2007coherent}.

%
% introduce machine learning
% 
In recent years, the field of machine learning and its sub-field, deep
learning, have been rapidly developing and have shown great potential in many
scientific fields~\cite{krizhevsky2012imagenet, NIPS2014_5423,
simonyan2014very, chen2014semantic, zeiler2014visualizing, szegedy2015going}.
For example, deep learning has been successfully applied to fields including
medical diagnosis~\cite{kononenko2001machine}, object
detection~\cite{redmon2016you}, image
recognition/processing/generation~\cite{he2016deep, krizhevsky2012imagenet,
zhang2016colorful, karpathy2015deep}, and  language
processing~\cite{lample2016neural}. In \ac{GW} astronomy, deep learning has
mostly been applied to the identifications of detector noise artefacts
(glitches)~\cite{mukund2017transient, zevin2017gravity, george2017deep} and the
detection of astrophysical signals~\cite{george2018deep, gabbard2018matching,
astone2018new} and their parameter estimation~\cm{add ref to Hunters latest
paper}. A \ac{CNN} is a deep learning algorithm that has the advantage of
capturing spatial and temporal features of the input data~\cm{this doesn't
really mean anything. Spatial refers to features in certain locations in images
and temporal refers to locations in a time series of in the time axis of a
spectrogram image}. Another advantage of the use of a \ac{CNN} in the detection
of a \ac{GW} signal is that a \ac{CNN} is relatively computationally cheap
compared to other more traditional methods. This is because the heavy
computational work is usually done during the training stage of a \ac{CNN}
prior to its actual application~\cite{goodfellow2016deep}.

%
% the main point of the paper
%
In this work, we demonstrate that a \ac{CNN} can be applied to the detection of
\acp{GW} from \acp{CCSN} and to the classification of their explosion
mechanisms.  To train our \ac{CNN}, we use simulated \ac{CCSN} waveforms from a
number of studies and simlate sources at a range of distances from $10$ to
$200$ kpc for two networks of four \ac{GW} detectors. The first network
consists of \ac{aLIGO}, \ac{AdVirgo} and KAGRA. For the second network, we
still include the detectors of \ac{AdVirgo} and KAGRA, but replace the two
detectors of \ac{aLIGO} with a modest set of planned upgrade version of them -
LIGO A+ in Hanford and Livingston~\cite{miller2015prospects, LIGOW}. We then
apply the \ac{CNN} to waveforms excluded from the training set to show that the
\ac{CNN} is not only able to identify the waveforms it familiarised itself with
during training, but also new waveforms~\cm{this is a strange statement. The
ability of the CNN to work on the training data is completely outweighed by its
perfromance on the independent testing set. The testing set perfroamnce is all
that you should mention here.}.

%
% paper structure
%
The remaining of this paper is structured as follows. In Section~\ref{sec:CNN},
we present a brief explanation of the concept of a \ac{CNN} as well as the
\ac{CNN} architecture  used for this study. In Section~\ref{sec:spwf}, we
introduce the waveforms we use for the training of the \ac{CNN} and  also
discuss the procedure with which we generated the input data for the \ac{CNN}.
The results are shown in Sections~\ref{sec:result} and~\ref{sec:unseen},
followed by our conclusions in Section~\ref{sec:conclusion}.
%\section{Supernova waveform}\label{sec:spwf}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Convolutional Neural network}\label{sec:CNN}

%
% describe the components of a CNN
%
A \ac{CNN} is a computational processing system that is composed of
interconnected layers of computational nodes~\cite{o2015introduction}. The
nodes are known as neurons and the outputs of which are processed with an
activation function. The activation function performs an elementwise nonlinear
operation to the output of the layer. In a standard \ac{CNN}, there are
commonly three types of layers: convolutional layers, max-pooling layers, and
fully connected layers~\cite{o2015introduction}. Convolutional layers perform
the mathematical operation of convolution between the weights of the layer's
neurons and the input to that layer. Max-pooling layers perform a down-sampling
process that reduces the size of the data by selecting the maximum of data
samples within fixed size bins. It can reduce the computational cost by
decreasing the number of trainable parameters (weights and biases) of the
\ac{CNN}.  Fully connected layers are layers that connect every neuron in its
layer to every neuron of its immediate previous and next layer. In the case of
detection and classification, the output layer is also a fully connected layer
with a softmax activation function with output used to represent the class
scores or probabilities.~\cm{even after editing, this isn't the clearest
description of a CNN. You are lacking an initial statement that says what a CNN
does - it takes in data and is able to classify it as one of the N types it has
learnt through training.}

%
% decribe the structure of the CNN
%
When multiple layers of these three types are stacked and connected one after
the other, a \ac{CNN} has been formed, where the output of each layer is the
input of the next layer. How these layers are connected in a \ac{CNN}, the
numbers of layers, their type, the number of neurons, the convolutional filter
size, max-pooling size, and the activation functions used, describes the
architecture of the \ac{CNN} (also known as hyper-parameters). The first layer
in a \ac{CNN}, also known as the input layer, is often a convolutional layer,
while the last layer or the output layer is often a fully connected layer with
an associated loss function~\cm{the loss isn't really associated with the
output layer. It is associated withe whole network.}. The architecture of the
remainder of the \ac{CNN} should depend on the specific task that it is being
trained to solve. An over-complicated model with too many trainable parameters
makes it easier to result in over-fitting (a procedure where the network
essentially memorises the training data but is unable to generalise to new
data). Alternatively, an overly simple architecture will struggle to capture
the features inherent to the input and not perform well in classification.
Automated schemes do exist for finding the optimal combination of
hyperparameters but in general (and in our case) the final architecture is
often obtained through rough trial and error, followed by fine tunning.

%
% describe training
%
During the training stage, the weights of the neurons in a \ac{CNN} are updated
using an algorithm called back-propagation~\cite{lecun1988theoretical}. The
output of the \ac{CNN} is used as an input to the loss-function~\cm{you need to
define what the loss function is supposed to do and how by minimising it you
make the network produce the right result. It isn't clear here.} and
back-propagation then computes the gradient of the loss-function with respect
to the traininable weights and biases within the network. A gradient descent
algorithm is then used to adjust the values of the weights and biases of the
neurons in each layer in order to iteratively minimise the loss-function.  When
the loss function is minimised and training is therefore complete, the \ac{CNN}
will be able to take in input in the form of new data and its output will best
represent the probability of that data belonging to each of the trained
classes. A well performing network will give high output probabilities to the
correct class in most cases. The process of achieving the minimisation of the
loss function during the training stage is the process whereby the machine is
"learning"..

%
% describe the specific architecture used here
% 
In this work, we employ a \ac{CNN} of $8$ convolutional layers, $3$ max-pooling
layers, and $3$ fully connected layers. The exact architecture of the \ac{CNN}
is shown in Table~\ref{table:architecture} and illustrated in
Fig.~\ref{fig:CNN}.

\begin{table}[]
\centering
\begin{threeparttable}
\caption{The architecture of the \ac{CNN}}
\label{table:architecture}
\begin{tabular}{ccccc}
\\
\toprule
Layer & Type      & Neurons  & Filter size & Act. Fun \\ \hline
1     & Conv      & 11        & 32          & Elu      \\ 
2     & Max-pool  &          & 8          &          \\
3     & Conv      & 11        & 8          & Elu      \\
4     & Max-pool  &          & 6           &          \\
5     & Conv      & 11       & 6           & Elu      \\
6     & Conv      & 11       & 4           & Elu      \\
7     & Conv      & 13       & 4           & Elu      \\
8     & Conv      & 13       & 4           & Elu      \\
9     & Conv      & 13       & 4           & Elu      \\
10    & Conv      & 13       & 4           & Elu      \\
11    & Max-pool  &          & 2           &          \\
12    & Fully-con & 64(50\%) &             & Elu      \\
13    & Fully-con & 32(50\%) &             & Elu      \\
14    & Fully-con & 3        &             & Softmax      \\ 
\hline
\hline
\end{tabular}
\begin{tablenotes}
\setlength\labelsep{0pt}
\normalfont{
\item The architecture of the \ac{CNN} used in this work for the purpose of
distinguishing supernova signals in backgound \ac{GW} detector noise vs detectr
noise alone. In the table, Conv means convolutional layer, Max-pool means
max-pooling layer, and Fully-con means fully-connected layer. Neuron and Act.
Fun indicate the number of neurons and the activation function used for the
layer respectively. The numbers in the bracket for the fully-connected layers
are the number used for drop-out~\cm{You haven't discussed drop-out yet. You
should mention it earlier when you introduce CNNs}.}
\end{tablenotes}
\end{threeparttable}
\end{table}

\begin{figure*}
\includegraphics[width=0.96\textwidth]{CNN2.png}
\caption{An illustration of the architecture of the \ac{CNN} used in this paper
for the detection and classification of \ac{CCSN} \ac{GW} signals in noisy data. The
\ac{CNN} consists of $8$ convolutional layers, $3$ max-pooling layers and $3$
fully connected layer including the output layer.  The input layer takes the
simulated time series of the detectors as input, feeding through the \ac{CNN}.
The \ac{CNN} will output three probabilities at the last layer.  The numbers
above or below each layer indicate the kernal size of the layer. For example,
the first convolutional layer has $11$ filters, each of which is $1$ by $32$ in
size~\cm{this is the only place that you describe the timeseries being the
input and what the network does in a single sentence. This text should appear
in the main bodty of the paper}.~\cm{Also, because I'm very picky about plots
and images - even though I made the image full page width, the timeseries plots
are TOO small and the axes are insanely small to read. You seem to have
strecthed curly bracket symbols to the point that they look super-wierd.}
\label{fig:CNN}}
\end{figure*}

%
% define the loss function 
%
The problem we are trying to solve is a problem of multi-class
classification, and hence the loss function employed for this work is the
categorical cross-entropy~\cite{abadi2016tensorflow}, defined as
%
\begin{equation}\label{eq:cce}
 L(y,\hat{y}) = -\sum^M_{j=1}\sum^C_{i=1}y_{ij}\log(\hat{y}_{ij}),
\end{equation}
%
where $C$ is the number of the classes and $M$ is the number of the training
samples~\cm{the loss is computed over batches so here $M$ should be the batch
size. However, you haven't introduced batches yet, which you need to do when
you're discussing training}. For the $j^{\text{th}}$ sample and the $i^{th}$
class, $y_{ij}$ is the correponding class value. It is equal to $1$ for the
true class and $0$ otherwise.  Similarly, $\hat{y}_{ij}$ is the predicted
probability from the \ac{CNN} for the $i^{th}$ class and the $j^{\text{th}}$
sample.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data}\label{sec:spwf}
%
% describe the classes and training, validation, and testing
%
We establish a \ac{CNN} for the purpose of distinguishing detector time series
among three classes, i.e., magnetorotational signals $+$ background noise,
neutrino-driven signals $+$ background noise, and pure background noise. For
this purpose, it is necessary to prepare training, validation and testing data
of these three classes. The training data is used for tuning the weights of
the neurons in the layers in the \ac{CNN}, while validating data is to verify
that the \ac{CNN} is learning the features inherent to the data and testing
data is to test the performance of the trained \ac{CNN}.

%
% describe the magneto waveforms
%
In our case, we define an input data sample as a set of simulated time series
stacked together as a $k \times p$ matrix where $k$ is the number of detectors
and $p$ the number of elements in the time series. To this end, we use
simulated waveforms taken from the literature. The magnetorotational \ac{CCSN}
signals are taken from~\cite{abdikamalov2014measuring,
dimmelmeier2008gravitational, richers2017equation}. The simulations
in~\cite{abdikamalov2014measuring} were focused on the dependence of the
waveforms on the angular momentum distribution of the progenitors and proivide
us with 92 waveforms. In~\cite{dimmelmeier2008gravitational}, 136 waveforms are
available from investigations into a variety of rotation rates and masses of
the progenitors. The simulations in~\cite{richers2017equation} covered a
parameter space of $18$ different equations of state and $98$ rotation profiles
for a progenitor of $12M_{\odot}$ providing a total of $1824$ waveforms. All
the simulations from the studies were based on 2D simulations of supernovae.

%
% describe the neutrino waveforms
%
For the neutrino-driven mechanism, we employ the waveforms
from~\cite{10.1093mnrasstz990, kuroda2017correlated, muller2012parametrized,
powell2019gravitational, radice2019characterizing, yakunin2015gravitational,
yakunin2017gravitational, ott2009gravitational, murphy2009model,
ott2013general}. The simulations in~\cite{10.1093mnrasstz990,
kuroda2017correlated, muller2012parametrized, powell2019gravitational,
radice2019characterizing, yakunin2017gravitational} were from 3D modelling of
the supernovae while the simulations in~\cite{yakunin2015gravitational,
ott2009gravitational, murphy2009model, ott2013general} were 2D. These
simulations cover a wide range of progenitor masses from $9\text{M}_\odot$ to
$60\text{M}_\odot$ and the specific progenitor masses of the simulations are
shown in Table~\ref{table:waveforms}. Examples of the simulated waveforms for
both mechanisms are shown in Figs.~\ref{fig:magwaveforms} and
\ref{fig:neuwaveforms}.

\begin{table}[]
\centering
\begin{threeparttable}
\caption{Waveforms}
\label{table:waveforms}
\begin{tabular}{rccc}
\toprule
    & Mechanism         & Mass ($\text{M}_\odot$)                         & No.  \\
\hline                         
Abdikamalov\cite{abdikamalov2014measuring} & M   & 12.0                          & 92   \\
Dimmelmeier\cite{dimmelmeier2008gravitational} & M   & 11.2,15.0,20.0,40.0        & 136  \\
Richers\cite{richers2017equation}     & M   & 12.0                          & 1824 \\
Andresen\cite{10.1093mnrasstz990}    & N   & 15.0                          & 6    \\
Kuroda\cite{kuroda2017correlated}      & N   & 11.2,15.0                    & 2    \\
Muller\cite{muller2012parametrized}      & N   & 15.0,20.0                    & 6    \\
Murphy\cite{murphy2009model}      & N   & 12.0,15.0,20.0,40.0        & 16   \\
Ott$_1$\cite{ott2009gravitational}       & N   & 15.0                          & 2    \\
Ott$_2$\cite{ott2013general}       & N   & 27.0                          & 8    \\
Powell\cite{powell2019gravitational}      & N   & 3.5,18.0                      & 2    \\
Radice\cite{radice2019characterizing}      & N   & 9,10,11,12,13,19,25,60 & 8    \\
Yakunin$_1$\cite{yakunin2015gravitational}   & N   & 12,15,20,25                & 4    \\
Yakunin$_2$\cite{yakunin2017gravitational}   & N   & 15                            & 1   \\
\hline
\hline
\end{tabular}
\begin{tablenotes}
\setlength\labelsep{0pt}
\normalfont{
\item The mass ranges and mechanisms of the progenitors of the simulated
waveforms used in this work.  The first column refers to the studies. Mechanism
indicates the explosion mechanism for the waveforms, with M for the
magnetorotational mechanism and N for the neutrino-driven mechanism.  Mass
refers the mass, in units of solar mass, of the progenitors in the simulations.
No. means the number of waveforms available from the study.} 
\end{tablenotes}
\end{threeparttable} 
\end{table}

%Anderson 2016 also does 3d, 20 mass, 11 mass
%kuroda 2017 3d, 15mass, 11mass
%powell 2018 3d 18mass, one model
%radics 9, 10, 11, 12, 13, 19, 25 and 60 mass 3d
%yakunin2015
%yakunin2017
%ott2009
%murphy2009model
%ott2013
%which cover both 2D simulations with masses of $15M_{\odot}$\cite{ott2009gravitational}, 
%and $12, 15, 20$ and $40M_{\odot}$\cite{murphy2009model}, and 3D simulations for a progenitor of $27M_{\odot}$\cite{ott2013general}.$26$

\begin{figure}
\includegraphics[width=0.5\textwidth]{mag_waveforms.png}
\caption{Examples of simulated waveforms used in the work for which the
explosion mechanism is magnetorotational mechanism.  The top panel shows a
waveform from~\cite{abdikamalov2014measuring}. The progenitor is
$12\text{M}_\odot$ with differential rotation parameter $A=300\text{km}$. The
initial angular velocity at the core $\Omega_c$ is $8.5\text{rads/s}$. The
middle panel shows a waveform from~\cite{dimmelmeier2008gravitational}. The
simulation corresponds to a progenitor of $15\text{M}_\odot$ with $A=50$km and
$\Omega_c = 1.91\text{rads/s}$.  The lower panel shows a waveform
from~\cite{richers2017equation} where the progenitor is $12\text{M}_\odot$, $A$
and $\Omega_c$ are $10000$km and $0.5\text{rad/s}$ respectively. In all panels,
only the $h_{+}$ polarisations are shown because the simulations are
axis-symmetric or 2D, and therefore described by only one polarisation.  The
sources are assumed to be at a distance of $10$ kpc. The time is not related to
the time at core bounce~\cm{then what is the time in reference to? Also, these
plots are too small and squashed. The text is certainly too small to be
readable and they should all have equal y and x-axis scales for comparison.}.
\label{fig:magwaveforms}} 
\end{figure}

\begin{figure}
\includegraphics[width=0.5\textwidth]{neu_waveforms.png}
\caption{Examples of simulated waveforms used in the work for which the explosion mechanism is neutrino-driven mechanism.
From the top to the bottom, the panels show waveforms from\cite{10.1093mnrasstz990, murphy2009model, radice2019characterizing}, with progenitors of masses equal to $15\text{M}_\odot$,
$15\text{M}_\odot$ and $60\text{M}_\odot$ respectively. 
The sources are assumed to be at $10$ kpc from earth. The time is not related to the time at core bounce.
\label{fig:neuwaveforms}}
\end{figure}
Introducing waveforms from a variety of studies essentially introduces a distribution of the waveforms that covers a larger parameter space, 
making the data samples harder for the \ac{CNN} to learn.
As a result, the performance of the trained \ac{CNN} on the testing samples 
may appear to be worse than that if the data were generated using only a subset of the waveforms.
However, a larger parameter space provides the advantage of forcing the \ac{CNN} to 
learn the features common among the waveforms, rather than a subset of the waveforms.
The trained network can thus perform better when facing waveforms with unexpected features, which is more likely in reality.

Since the waveforms are generated at various distances, sampling rates and durations, 
it is necessary to normalise the waveforms before they can be used for the generation of the time series.
To do this, we first scale the amplitudes of the waveforms by moving the sources to $10$ kpc from earth.
We then ensure that the sampling rate are identical for all the waveforms by down sampling 
them to a pre-selected sampling rate (e.g., $4096$Hz).
The longest duration $\tau$ among the waveforms is then identified and each of the remaining waveforms is padded with zeros
to this duration.
To introduce as less artefact as possible, a high pass filter with a low cut-off frequency equal to $11$Hz 
and a tukey window ($\alpha = 0.08$) are applied prior to the zero padding. 
To balance the difference in the number of waveforms 
between the two mechanisms in the final data set, $36$ copies of each neutrino-driven waveform are duplicated.
After this procedure, the simulated 
waveforms are then $\bold{S}(t) = \{\bold{s}_{1}(t), \bold{s}_{2}(t), ..., \bold{s}_{m}(t)\}$, where $m$ is the number of the waveforms and 
$\bold{s}_{q}$ is the $q^{\text{th}}$ waveform, defined by,
\begin{equation}\label{eq:ht}
 \bold{s}_q(t) = \left(\begin{array}{c}
                        h_q^+(t) \\
                        h_q^\times(t)
                       \end{array}\right),
\end{equation}
where $h_q^+(t)$ and $h_q^\times(t)$ are the two polarisations of the waveform and $t$ the time. 
As mentioned, some of the waveforms used in the work are generated with the simulations being axis-symmetric, the \acp{GW} are entirely described by one polarisation $h_q^+(t)$. 
The corresponding $h_q^\times(t)$ for these waveforms are a zero vector. 
In this work, we perform simulations for distances equal to $10$, $20$, $30$, $40$, $50$, $60$, $80$, $100$, $150$ and $200$ kpc.
If for a training session, the distance $d_\text{L}$ interested is not $10$ kpc, the following relationship is employed to scale the amplitudes of the waveforms,
\begin{equation}\label{eq:sap}
 \bold{s}^{d_\text{L}}_q(t) =  \frac{10\text{kpc}\times\bold{s}_q(t)}{d_\text{L}},
\end{equation}
where $\bold{s}^{d_\text{L}}_q(t)$ is the waveform if the source is at $d_\text{L}$ from earth.

The next step is to generate simulated time series for the \ac{GW} detectors in a network using $\bold{S}(t)$.
%The input data consist of simulated time-series from a network of \ac{GW} detectors.
Since the purpose of building a \ac{CNN} is to categorise an input data into three exclusive classes,
in total three types of time-series are generated.
For the time series containing either a magnetorotational signal or a neutrino-driven signal, 
we start by selecting a waveform $\bold{s}_q$ from $\bold{S}(t)$.
A random location of (right ascension, declination) $=(\alpha, \delta)$ in the sky is selected from a uniform distribution on $\alpha$ and a uniform distribution on the sine of $\delta$.
The corresponding antenna patterns $\bold{F}(\alpha, \delta, t)$, given by the following equation,
\begin{equation}\label{eq:ap}
\bold{F}(\alpha, \delta, t)= \left(\begin{array}{cc}
f_1^+(\alpha, \delta, t)~~f_1^\times(\alpha, \delta, t) \\
\vdots~~~~~~~~~~~~\vdots \\
f_k^+(\alpha, \delta, t)~~f_k^\times(\alpha, \delta, t) \\
\end{array}\right),
\end{equation}
is computed, 
where $f^+(\alpha, \delta, t)$ and $f^{\times}(\alpha, \delta, t)$ are the antenna pattern functions for the two polarizations and $k$ is the number of detectors as defined above.
The relative delays in the arrival times for the location are also computed and applied to the selected waveform.
The delay in arrival time between a detector and the center of the earth is given by,
\begin{equation}\label{eq:tdetector}
%
\Delta t = \frac{\bold{n} \cdot \bold{r}}{c},
%
\end{equation}
%
where $\bold{n}$ is the propagation direction of the \ac{GW}, $c$ the speed of light, and $\bold{r}$ the location vector of the detector
relative to the center of the Earth. 
The resulting signal $ \bold{h}_j(\alpha, \delta, t)$ as received at the detectors is then given by the following equation,
\begin{equation}\label{eq:ht2}
 \bold{h}_j(\alpha, \delta, t) = \bold{F}(\alpha, \delta, t) \times \bold{s}_q(t),
\end{equation}
where the time delays $\Delta t$ are absorbed into the corresponding notations and the subscript $j$ is defined in Eq.\ref{eq:cce}.
Next, we generate independent Gaussian noise $\bold{N}_j(t) = \{\bold{n}_{1j}(t), \bold{n}_{2j}(t), ..., \bold{n}_{kj}(t)\}'$ 
for each detector in the network using their respective power spectral densities.  
In the above notation, the symbol $'$ indicates transpose. 
%It is worth mentioning that although in reality, 
%noise at \ac{GW} detectors may not be entirely Gaussian and contaminated with glitches,
%the application of \ac{CNN} is actually not limited to time series where the noise is Gaussian.
%Although  we will focus on Gaussian noise in this work, 
%a \ac{CNN} can in fact be trained to identify the presence of a glitch in the data \cite{mukund2017transient, zevin2017gravity, george2017deep, gabbard2018matching}.
The duration of the generated noise is $1.6$ times longer than $ \bold{h}_j(\alpha, \delta, t)$. 
A random number $x$ will then be generated determining where in the generated noise the signal will be placed, as given by,
\begin{equation}
\bold{d}_j(\alpha, \delta, t, x)=
\begin{cases}
\bold{N}_j(t)~~~~~~~~~~~~~~~~~t < x; \\
\bold{N}_j(t) + \bold{h}_j(\alpha, \delta, t)~x \leq t \leq \tau + x; \\
\bold{N}_j(t)~~~~~~~~~~~~~~~~~t > \tau + x. \\
\end{cases}
\end{equation}
This is to avoid the possibility that the \ac{CNN} learns human artefact instead of common features of the waveforms by having the signals always starting at the same place.
For each time series of the background noise class, independently simulated Gaussian background noise of the same duration as that of the other classes 
are generated for each detectors in the network.
This means a data sample for this class is defined as,
\begin{equation}\label{eq:noise}
 \bold{d}_j(t) = \bold{N}_j(t).
\end{equation}

Since training a \ac{CNN} of a given structure to its maximum capacity requires many data samples, we augment our data set by iterating over $\bold{S}(t)$ until 
for each waveform in $\bold{S}(t)$, we have $31$ data samples $\bold{d}$ generated using the procedure described above.
This step is essential for the \ac{CNN} to learn the features of the waveforms and to identify them effectively under different noise scenarios.
After that, we generate independent noise realisations for the background noise class.
The entire data set $\bold{D}$ defined below, has roughly $l = 1.8 \times 10^{5}$ data samples, where each class has approximately $6\times10^{4}$ data samples.
\begin{equation}\label{eq:finaldata}
\bold{D}(\boldsymbol{\alpha}, \boldsymbol{\delta}, \boldsymbol{\delta_t}, \bold{t}, \bold{x}) =  \left(\begin{array}{c}
\bold{d}_1(\alpha, \delta, \delta_t, t, x) \\
\vdots \\
\bold{d}_o(\alpha, \delta, \delta_t, t, x) \\
\vdots \\
\bold{d}_l(t) \\
\end{array}\right).
\end{equation}
In the above equation, the subscript $o$ is the sum of the data samples in which there is a \ac{GW} signal in the time series. 
The final step is to whiten $\bold{D}$ using the power spectral densities.
In Figure \ref{fig:sample}, a representative example  of time series for one detector is shown. 
\begin{figure}
\includegraphics[width=0.5\textwidth]{datasample.png}
\caption{Representative example of the simulated time series used to train/validate/test the \ac{CNN}.
The blue shows a whitened time series with a signal buried in Gaussian noise of unit variance.
The red shows the same signal free of noise and whitened. 
\label{fig:sample}}
\end{figure}
A data sample consists of such time series of the same signal or background noise from all the detectors in the network.
The data samples are then split, with $1\times10^{4}$ samples being randomly taken for validation, $10\%$ of the rest for testing and $90\%$ for training.
When the training is finished for a distance, 
the above described procedure will be repeated for another distance until the training
for all distances have been carried out for a network.
The entire procedure is then repeated for another network of \ac{GW} detectors.
As mentioned previously, the \ac{CNN} will be trained for two networks of \ac{GW} detectors.
We present the networks in Table \ref{table:network}. For the remaining of the paper, we will use their acronyms to refer to the networks.
\begin{table}[]
\centering
\begin{threeparttable}
\caption{Detector networks}
\label{table:network}
\begin{tabular}{ccc}
\toprule
Network            & Detectors                & Acronym                 \\
\hline
\multirow{4}{*}{1} & \ac{aLIGO} H*    & \multirow{4}{*}{HLVK}   \\
                   & \ac{aLIGO} L* &                         \\
                   & \ac{AdVirgo}           &                         \\
                   & KAGRA                    &                         \\
                   \hline
\multirow{4}{*}{2} & LIGO A+ H          & \multirow{4}{*}{H+L+VK} \\
                   & LIGO A+ L       &                         \\
                   & \ac{AdVirgo}           &                         \\
                   & KAGRA                    &                         \\
\hline
\hline
\end{tabular}
\begin{tablenotes}
\setlength\labelsep{0pt}
\normalfont{
\item The networks of detectors used in this work.\\
* H refers to the detector in Hanford and L the detector in Livingston.}
\end{tablenotes}
\end{threeparttable}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Result and discussion}\label{sec:result}
After the \ac{CNN} is trained, we can estimate its performance using the testing samples.
The results are presented in this section.
One of the most used and convenient ways to determine the classifying performance of a model is to plot the \ac{ROC} curve.
A \ac{ROC} shows the performance of a classifying model by showing the \ac{TAP} at a given \ac{FAP}. 
Since \ac{ROC} is usually plotted for model distinguishing two classes, 
for a multi-class classification problem, 
a \ac{ROC} for a class should be viewed as the class versus the others.
This means in this context, \ac{FAP} means the fraction of samples from other 
classes misidentified as a sample from the class the \ac{ROC} is associated with.
\ac{TAP} is identical to that of two-class classification problem and indicates the fraction of samples correctly identified.
For a given \ac{FAP}, a model with a higher \ac{TAP} is considered more capable than a model with a lower \ac{TAP}.
In Figure \ref{fig:ROClog}, we show the \acp{ROC} for both the mechanisms and \ac{GW} detector networks tested in this work. 
For simplicity, we show only the results for three distances, namely, $20$, $60$ and $100$ kpc.
\begin{figure*}
     \begin{center}
%
        \subfigure[]{%
            \label{fig:ROClog1}%
            \includegraphics[width=0.45\textwidth]{ROC_curves_neutrino.png}%
        }\quad
        \subfigure[]{%
            \label{fig:ROClog2}
            \includegraphics[width=0.45\textwidth]{ROC_curves_mag.png}%
        }%
%
    \end{center}
    \caption{ROC curves showing the classification performance of the \ac{CNN} for \ac{CCSN} signals of different explosion mechanisms at three distances, $20$, $60$, and $100$kpc.
The left panel shows the results for the neutrino-driven mechanism, while the panel on the right shows those for the magnetorotational mechanism.
In both panels, the solid lines are for the network H+L+VK, while the dashed lines are for HLVK.
\label{fig:ROClog}}%
\end{figure*}
For all the distances tested, the \ac{CNN} achieves a higher \ac{TAP} for any given \acp{FAP} 
for magnetorotational than neutrino-driven signals.
That is not surprising as the amplitudes for magnetorotational signals are higher than that of neutrino-driven signals.

We can also show the classification efficiency of the \ac{CNN}  as a function of distance.
This is done by fixing the \ac{FAP} and plotting the \ac{TAP}. 
The results for three chosen \acp{FAP} are shown in Figure \ref{fig:eff}.
\begin{figure*}
     \begin{center}
%
        \subfigure[]{%
            \label{fig:eff1}%
            \includegraphics[width=0.45\textwidth]{Neutrino_efficiency.png}%
        }\quad
        \subfigure[]{%
            \label{fig:eff2}
            \includegraphics[width=0.45\textwidth]{Magnetarotational_efficiency.png}%
        }%
%
    \end{center}
    \caption{Efficiency curves showing the classification ability of the \ac{CNN} as a function of distance for both mechanisms and networks.
The left panel shows the results for the neutrino-driven mechanism, while the right shows those for the magnetorotational mechanism. 
In both panels, the solid lines show the results for the network H+L+VK, and the dashed lines for HLVK.
Three \acp{FAP} are chosen, i.e.: blue for \ac{FAP} $=0.1$, green for \ac{FAP} $=0.01$, red for \ac{FAP} $=0.001$.
\label{fig:eff}}%
\end{figure*}
In this figure, a similar trend is seen that magnetorotational signals are easier for the \ac{CNN} to identify than neutrino-driven signals at any distance. 
For magnetorotational signals from sources located at $50$ kpc and a \ac{FAP} of $0.1$, 
the \ac{CNN} achieves a \ac{TAP} of $94\%$ and $82\%$ for the networks H+L+VK and HLVK respectively.
At more restrict \acp{FAP} such as $0.001$, the \ac{CNN} still achieves \acp{TAP} $86\%$ and $68\%$ for sources at the same distance for the two networks respectively.
For sources that are located at a slightly further distance of $60$ kpc, both networks achieve a \ac{TAP} of close to or larger than $80\%$ at a \ac{FAP} of $0.1$.  
Such a range is well into the Large and Small Magellanic Clouds and covers the satellite galaxies in between\cite{karachentsev2004catalog, belokurov2007cats}.
For sources at $100$ kpc, the \acp{TAP} are close to or larger than $60\%$ for H+L+VK for all chosen \acp{FAP}, and is $73\%$ if the \ac{FAP} is $0.1$.  
Even for sources at $150$ and $200$ kpc, the \acp{TAP} are $54\%$ and $38\%$ respectively for the same \ac{FAP}, indicating that with such a \ac{GW} network,
it is possible to detect magnetorotational \ac{CCSN} signals out to such a distance.
On the other hand, it is more difficult for the \ac{CNN} to detect and classify the neutrino-driven signals, due to their weaker amplitudes.
Nonetheless, for sources at $10$ kpc,
the \ac{CNN} achieves a \ac{TAP} of $76\%$ and $55\%$ for H+L+VK and HLVK respectively if the \ac{FAP} is $0.1$.
This means that the \ac{GW} from a Galactic \ac{CCSN} signal is possible to be detected and classified with either of these networks.
\begin{figure*}
     \begin{center}
%
        \subfigure[]{%
            \label{fig:ROCfixed1}%
            \includegraphics[width=0.45\textwidth]{ROC_fixed_neutrino.png}%
        }\quad
        \subfigure[]{%
            \label{fig:ROCfixed2}
            \includegraphics[width=0.45\textwidth]{ROC_fixed_mag.png}%
        }%
%
    \end{center}
    \caption{Efficiency curves showing the ability of the \ac{CNN} in distinguishing input data with a fixed decision threshold(0.5), and their corresponding \acp{FAP}.
    The left panel shows the results for the neutrino-driven mechanism, and the right shows that for the magnetorotational mechanism.
    In both panels, the solid lines show the \acp{TAP} and \acp{FAP} for the network H+L+VK, and the dashed lines show those for HLVK.  
\label{fig:ROCfixed}}%
\end{figure*}

In practice, since the output of a \ac{CNN} is probabilities indicating how likely the input belongs to each of the classes, 
it may be desired to set a threshold on which the decision whether the input belong to a class is made.
For example, an input will be classified into a class if the corresponding probability is larger than the pre-selected threshold.
In such a scenario, the \ac{FAP} and \ac{TAP} would be affected by the choice of the threshold.
We show such a result in Figure \ref{fig:ROCfixed}, where we employed a threshold of $0.5$. 
For H+L+VK and magnetorotational signals at $10$ kpc, the \ac{TAP} is $99\%$. 
If the distance is extended to $80$ kpc, the \ac{TAP} is still close to $80\%$.
For the largest distances tested in this work, $150$ and $200$ kpc, the \acp{TAP} are $43\%$ and $29\%$ respectively.
For HLVK, the \ac{TAP} is $98\%$ and $49\%$ for sources at $10$ and $80$ kpc respectively.
%The \acp{TAP} decrease to $10\%$ and $9\%$ for sources at $150$ and $200$ kpc.
Throughout the distances, the \ac{CNN} maintains a \ac{FAP} no larger than $0.04$ for both networks.
For neutrino-driven signals, it is shown that H+L+VK has a \acp{TAP} of $66\%$ at $10$ kpc while it is $55\%$ at $10$ kpc for HLVK.
Both of the networks have \acp{FAP} close to or less than $0.2$.
It should be noted that the results presented here are averaged over all the data samples in the testing samples.
The performance on any individual waveform may vary depending on the morphology and the amplitudes of the waveform.

%Also shown in this figure is a curve showing the \ac{TAP} for all the classes, defined by,
%\begin{equation}\label{eq:TAPforall}
%\text{TAP}_{\text{all}} = \frac{\text{No. of correctly identified test input}}{\text{No. of test input}}.
%\end{equation}
%This curve provides an overall estimate of the fraction of input that the \ac{CNN} will correctly classify.
%Unsurprisingly, this curve resides between the lines for magnetorotational and neutrino-driven signals.

\section{Waveforms Excluded from Training}\label{sec:unseen}
The above section proves that using a \ac{CNN}, it is possible to detect and classify \ac{CCSN} waveforms 
if the waveforms have been used to train the \ac{CNN}. 
However, in reality, it is likely that the \acp{GW} from a \ac{CCSN} may only be partially similar to the simulated waveforms, 
while having some other features that are different or even unexpected. 
A \ac{CNN} that is only capable of recognising waveforms with which it is familiar may prove to be not entirely applicable.
Therefore, we apply our trained \ac{CNN} to waveforms from other studies that were not used during the training.
Similar to what was done in \cite{roma2019astrophysics}, 
we take the waveforms $\text{R3E1AC}$ and $\text{R4E1FC\_L}$ from \cite{scheidegger2010influence} as the test waveforms for the magnetorotational mechanism.
The test waveforms for the neutrino-driven mechanisms are $\text{s}20$ from \cite{andresen2017gravitational} and $\text{SFHx}$ from \cite{kuroda2016new}.
The waveforms are shown in Figure \ref{fig:Extratestwaveform}.
The test waveforms in this section should not be confused with the testing samples in section \ref{sec:result}, 
which are randomly drawn from the same distribution as the training samples.
\begin{figure}
     \begin{center}
%
        \subfigure[]{%
            \label{fig:ex_waveform_neu}%
            \includegraphics[width=0.45\textwidth]{ex_waveform_neu.png}%
        }\quad
        \subfigure[]{%
            \label{fig:ex_waveform_mag}
            \includegraphics[width=0.45\textwidth]{ex_waveform_mag.png}%
        }%
%
    \end{center}
    \caption{The waveforms excluded from the training session of the \ac{CNN} and used as an extra test for the performance of the \ac{CNN}.
    From the top to the bottom, the waveforms shown are $\text{SFHx}$, $\text{s}20$(neutrino-driven mechanism), 
    $\text{R3E1AC}$ and $\text{R4E1FC\_L}$(magnetorotational mechanism), respectively. 
    The time in the x-axis is not related to core bounce.
\label{fig:Extratestwaveform}}%
\end{figure}
\begin{figure}
     \begin{center}
%
        \subfigure[]{%
            \label{fig:extra_waveform_test_neu}%
            \includegraphics[width=0.45\textwidth]{extra_neu_waveform_test.png}%
        }\quad
        \subfigure[]{%
            \label{fig:extra_waveform_test_mag}
            \includegraphics[width=0.45\textwidth]{extra_mag_waveform_test.png}%
        }%
%
    \end{center}
    \caption{Efficiency curves showing the ability of the \ac{CNN} in distinguishing input data.
    The upper panel shows the results for the the waveforms $\text{SFHx}$ and $\text{s}20$ (neutrino-driven mechanism), 
    and the lower panel shows that for the waveforms $\text{R3E1AC}$ and $\text{R4E1FC\_L}$ (the magnetorotational mechanism).
    In both panels, the solid lines show the \acp{TAP} and \acp{FAP} for H+L+VK, and the dashed lines show those for HLVK.  
    The waveforms have not been used for the training of the \ac{CNN}.
\label{fig:Extratest}}%
\end{figure}
Using the procedure described in section \ref{sec:spwf}, we generate $1200$ injections for each distance and each class has around 400 injections.
The performance of the \ac{CNN} on the test waveforms are shown in Figure \ref{fig:Extratest}.
For waveforms $\text{R3E1AC}$ and $\text{R4E1FC\_L}$ at $50$kpc and a \ac{FAP} of $0.1$, the \ac{CNN} achieves a \ac{TAP} of $87\%$ with H+L+VK and $59\%$ with HLVK respectively.
At $60$kpc, the \acp{TAP} drop slightly to $83\%$ and $52\%$ respectively.
This indicates that with the \ac{CNN}, it is possible to detect these waveforms from sources that are well into the Large and Small Magellanic Clouds.
For waveforms $\text{s}20$ and $\text{SFHx}$, the \acp{TAP} are $93\%$ and $70\%$ for the two networks if the sources are at $10$kpc and the \ac{FAP} is $0.1$. 
The efficiency for the waveforms $\text{s}20$ and $\text{SFHx}$ is noticeably higher than that for the neutrino-driven mechanism in Figure \ref{fig:eff} for the same distance.
This is because as mentioned before, the detection and classification efficiency in Figure \ref{fig:eff} are the results averaged over the testing samples.
It is possible that individual waveforms may be easier for the \ac{CNN} to detect. The results indicate that the \ac{CNN} has generalised the features of the explosion mechanisms.
\section{Conclusion}\label{sec:conclusion}
We have demonstrated a \ac{CNN} can be applied for the purpose of distinguishing \ac{GW} detector time series 
among pure background noise and \ac{CCSN} explosion mechanisms. 
We trained the \ac{CNN} using $1.8\times10^{5}$ samples of simulated time series for each distance in a number of distances from $10$kpc to $200$kpc.
The data samples for each classes consisted of approximately $4\times10^5$ samples.

We have shown that with a network of HLVK, when the \ac{FAP} was $0.1$, once trained, a \ac{CNN} could achieve a \ac{TAP} close to $80\%$
for magnetorotational signals from sources at $60$ kpc. 
Using a network of H+L+VK, we showed that the \ac{TAP} could be even more promising at $91\%$. Both the Large and Small Magellanic Clouds are within this distance.
If the distance is extended to $150$ or $200$ kpc, a \ac{TAP} of $54\%$ or $38\%$ 
respectively are still achievable for H+L+VK, indicating the possibility of detections within such distances.

For the neutrino-driven mechanism, the weaker amplitudes of the waveforms result in lower \acp{TAP} at the same distances.
For sources at $10$kpc, the trained \ac{CNN}, with a \ac{FAP} of $0.1$, achieved a \ac{TAP} of $55\%$ and $76\%$ for HLVK and H+L+VK respectively.
This indicates a Galactic \ac{CCSN} event is likely to be detectable.

We used four waveforms that were not used for the training of the \ac{CNN} to test the performance of the \ac{CNN} in a more realistic situation.
We found that for waveforms $\text{R3E1AC}$ and $\text{R4E1FC\_L}$ from sources at $60$kpc, the \acp{TAP} are $83\%$ and $52\%$ for H+L+VK and HLVK respectively.
For waveforms $\text{s}20$ and $\text{SFHx}$, the \acp{TAP} are $93\%$ and $70\%$ for the two networks respectively.
The results prove the possibility of using \ac{CNN} for the detection and classification of \ac{CCSN} \ac{GW} signals.
\\
\section*{ACKNOWLEDGEMENTS}
I.S.H. and C.M. are supported by
the Science and Technology Research Council (grant No.
ST/L000946/1) and the European Cooperation in
Science and Technology (COST) action CA17137. 
\bibliography{main}

\end{document}