### 2020-04-28 Response_to_PRD_reviewers.txt
Fixed minor typos in the referee response

We would like to express our gratitude to the reviewers for reading 
through the paper thoroughly and their positive and constructive 
comments. Below are our response to the comments.

Report of Referee A 
------------------------------------------------------------------------------
The manuscript aims at the classification of gravitational wave
signals buried in noise through a convolutional neural network that
takes time series data from gravitational wave detectors as input and
works independently. The authors took into account simulated
waveforms, generated at various distances, for which the explosion
mechanism is modelled by the neutrino-driven mechanism and the
magneto-rotational mechanism. After the training of the convolutional
neural network, the authors estimated the performance of the algorithm
by constructing the receiver operator characteristic curves. The
waveforms excluded from the training session of the convolutional
neural network are used as an extra test for the performance of the
trained CNN. The manuscript is well organised and clearly written, it
offers a complete overview of the state of art and it contains a
significant step forward for the classification of supernova
gravitational waves signals.

Response: Thank you so much for the very positive comments. 
          They are much appreciated.
------------------------------------------------------------------------------

Report of Referee B
------------------------------------------------------------------------------
The authors present a novel analysis pipeline (at the heart of which
is a convolutional neural network - CNN) for the detection and
classification of gravitational waves (GWs) from core-collapse
supernovae (CCSNe). Employing said pipeline, the authors suggest GWs
from Galactic CCSNe are likely to be detected with even the current
generation of ground-based GW detectors. If robust, this result would
provide an improvement in sensitivity by a factor of a few over
similar studies using excess-power methods, which is exciting and much
needed.

The particulars of the conditions for which the flagship paper results
are quoted, however, cause concern. A Galactic CCSN is a once in a
lifetime event, and the first "detection" of GWs from stellar collapse
would certainly not be claimed at any less than 5 sigma. As a
consequence, any results presented at FAP=1% or FAP=10% on
detectability seem a little spurious.

Going through the paper and finding the equivalent
"detection/classification distances" at FAP=0.1% (the most realistic
case presented, and a little over 3 sigma):

* [Fig. 5a] At 20kpc, TAP_HLVK ~ 1%, and TAP_H+L+VK ~ 8%

* [Fig. 5b] At 20kpc,   TAP_HLVK ~ 92% / TAP_H+L+VK ~ 95%
                At 60kpc,   TAP_HLVK ~ 55% / TAP_H+L+VK ~ 80%
                At 100kpc, TAP_HLVK ~ 16% / TAP_H+L+VK ~ 16%

* [Fig. 6a] At 10kpc, TAP_HLVK ~ 12% / TAP_H+L+VK ~ 41%

* [Fig. 6b] At 10kpc, TAP_HLVK ~ 96% / TAP_H+L+VK ~ 98%

* [Fig. 9a] At 10kpc, TAP_HLVK ~ 12% / TAP_H+L+VK ~ 41%

* [Fig. 9b] At 10kpc, TAP_HLVK ~ 96% / TAP_H+L+VK ~ 97%

These numbers (and the above commentary) suggest that a different
region of parameter space might be more appropriate to study here.
Assuming TAP>50% and FAP<0.3% as a (generous) lower limit on realistic
detection claims, it would be instructive to instead show:

* [Fig. 5a] TAP_HLVK vs FAP_HLVK and TAP_H+L+VK vs FAP_H+L+VK between
0.001% < FAP < 1% at 0.2kpc (Betelgeuse), 2kpc, and 10kpc (typical
"Galactic SN" distance)

* [Fig. 5b] TAP_HLVK vs FAP_HLVK and TAP_H+L+VK vs FAP_H+L+VK between
0.001% < FAP < 1% at 2kpc, 20kpc, 60kpc

* [Fig. 6a] TAP_HLVK and TAP_H+L+VK as a function of distance between
0.5kpc - 20kpc at FAP=0.1%, 0.01%, and 0.001%

* [Fig. 6b] TAP_HLVK and TAP_H+L+VK as a function of distance between
0.5kpc - 200kpc at FAP=0.1%, 0.01%, and 0.001%

* [Fig. 9a] TAP_HLVK and TAP_H+L+VK as a function of distance between
0.5kpc - 20kpc at FAP=0.1%, 0.01%, and 0.001%

* [Fig. 9b] TAP_HLVK and TAP_H+L+VK as a function of distance between
0.5kpc - 200kpc at FAP=0.1%, 0.01%, and 0.001%

Looking at Fig. 7, however, I question if it is possible to achieve
such low FAPs with this pipeline, particularly for the neutrino
mechanism waveforms. It would be useful to remake this figure at
distances for which FAP << 0.3% is possible. If not possible, it may
be beneficial to remove the figure. If not removed, the caption and
text should be updated to note the large FAPs used.

Response: We have tested again the trained CNN with 1.08^6 test samples 
          for each distance so that we can achieved a FAP of 0.1%, 0.01% 
          and 0.001%.

          We appreciate that re-running the simulations at closer
          distances is advised, but doing so would be very 
          computationally expensive.

          The changes in response to this comment includes: 
          1. The plots in the section of results (section IV and section V) 
             have been updated.  The text in these two sections have been 
             modified to reflect these updated results.

          2. Only TAP values at FAP equal to 0.1% are quoted in the      
             abstract and the conclusion. No detection claim is made for 
             FAP higher than 0.1%. The text in the abstract and the 
             conclusion has also been modified to accommodate the changes.

          3. We retracted claims regarding the ability of the 
             networks in detecting Galactic events where the explosion 
             mechanism is the neutrino-driven mechanism.

          4. The number of training/test samples in the last paragraph of 
             section III has been updated to be consistent with 
             this change.

          5. We would also like to highlight the fact that the plots
             previously submitted as Fig. 9a and Fig. 9b were actually 
             duplicates of previous Fig. 6a and Fig. 6b by accident. 
             Although this mistake was naturally corrected when we updated
             the results, we would still like to bring this mistake to the
             reviewers' attention.

Overall, the study is a novel and interesting work. It is crucial,
however, to put the authors' pipeline through more rigorous testing
(i.e. at realistic FAPs, perhaps using realistic noise) before stating
its ability to detect GWs from Galactic CCSNe. I would suggest framing
instead the paper as a follow-up pipeline, but take note that the
authors explicitly distinguish their pipeline from that shown in a
similar work by Astone+2018 as designed to take time series directly
from the detectors and find signals, rather than analyzing triggers
found by some external detection pipeline. Given this, it seems
appropriate for appreciable corrections to the text and results to be
requested.

Response: We appreciate this comment and admit that testing/training 
          the CNN using realistic noise would indeed show a more 
          realistic results.
          However, such an extension of the current work would require 
          a whole new set of treatments that we feel is beyond the current scope
          of this work. 
          We have therefore updated our results with more realistic 
          values of FAP and leave tests using realistic noise for future 
          work.

Below we list specific comments by section, denoting the
relevant paragraph number by p.

I: Introduction

p1: LIGO, Virgo, and KAGRA are not introduced or cited here. Suggest
to move discussion of second generation detectors (and references)
from start of p2 here. Aside - has KAGRA now begun operations? If so,
it may be worth updating the text here.

Response: We have updated the text accordingly.

p2: Specifically, the sentence "The nuclear equation of state will
then stiffens by the strong nuclear force above the nuclear density
and stops the core collapse." doesn't really make sense. The equation
of state stiffens (and subsequently stops the core collapsing) as the
core density exceeds nuclear density because the strong nuclear force
becomes repulsive at very small inter-nucleon distances.

Response: Thanks for pointing this out. We have rewritten the sentence 
          to read "The nuclear equation of state will then stiffens by 
          the strong nuclear force as the core exceeds nuclear density
          and stops the core collapse."

II: Convolutional Neural Network & Fig. 1 The technical description of
the CNN pipeline is thorough and instructive - nice work!

Response: Thank you so much.

Tab. 1 caption: typo at end of first sentence, detectr --> detector

Response: This has been corrected.

III: Data

p3: For the 3D simulations, what is the observer angle for the h+,hx
used? As the GW signal for CCSNe varies with orientation of the source
with respect to the observer, it may be useful to add a distribution
of observer angles for each waveform model to the training set, so as
to account for this variance. This could also help remedy the issue
noted in p6 of small sample size of neutrino mechanism waveforms
compared to magnetorotational mechanism waveforms.

Response: The observer angles for all the studies of 3D simulations have
          been added to the text. We appreciate that our data could be 
          augmented by varying the observer angles in the data set. 
          However, we noted that while some of the used studies published 
          the information or parameters to generate waveforms at random 
          observer angles, the same information for other studies is not
          readily available. We therefore decided to use only the published
          waveforms.

Fig. 2 & Fig. 3: What are the observer angles for these waveforms? As
an aside, it could be useful to combine these plots into a 4 panel
figure.

Response: the observer angles for these figures have been added in 
          the caption.

Note: Appreciated the technical detail included on signal injection
methodology - very refreshing to see, and crucial for reproducibility.

Response: Thank you so much.

IV: Results and Discussion

I suggest the results are rerun using the parameters highlighted
above, and the text subsequently rewritten to reflect the updated
results. At the very least, statements on detecting GWs from a
Galactic CCSN should be removed (unless results with FAP < 0.3% show
this is the case), and a clear note/disclaimer on the high FAPs used
must be added.

Note: Figs. 5 and 6 are clear and well presented - very instructive.

Response: As stated above, we have updated section IV and section V.

V: Waveforms Excluded From Training

As for the results section, I suggest reruns using parameters
highlighted above, and the text subsequently rewritten to reflect the
updated results. Similarly to before, at the very least a disclaimer
on the FAPs must be added, and detection statements for FAP>0.3%
removed.

Response: As stated above, we again tested the trained CNN with 
          1.08 * 10^6 test samples to achieve FAP values of 0.1%, 0.01%, 
          0.001% as was done for section IV. The text and plots 
          are also updated to reflect this change. 
          Claims regarding the ability of the networks in detecting 
          Galactic CCSN signals if the explosion mechanism is the 
          neutrino driven mechanism are withdrawn.

Fig. 8: As for Figs. 2 and 3, what are the observer angles for these
waveforms. As it is noted that h+ >> hx in the caption, I suspect the
waveform is for a polar observer, but this should be confirmed and
noted in the text.

Response: The waveforms shown in Fig. 8a and Fig. 8b are indeed for a 
          polar observer. This as well as that for Fig. 8c and Fig. 8d
          have been added in the caption.

VI: Conclusion

As previously noted, only results at FAP=10% are mentioned. As it is
not realistic to claim detection at such high FAP, the distances
presented are meaningless. This should be rewritten, pointing instead
to the results at FAP=0.1%. If Galactic/Magellanic Cloud detections
are no longer feasible, the abstract (and other relevant sections)
should be rewritten to reflect that.

Response: We updated the text to be consistent with the uses of more realistic 
          FAP and retracted claims of detection of Galactic event with HLVK.
------------------------------------------------------------------------------